---
title: "FCAT Ant Survey"
author: "Bolívar Aponte Rolón"
date: "2023-07-27"
output: html_document
---
# Packages
```{r}
library(vegan)
library(dplyr)
library(hillR)
library(GGally)
library(iNEXT)
library(ggplot2)
library(ggpubr)
library(tidyverse)
library(reshape2)
library(gridExtra)
library(lme4)
library(dplyr)
library(betapart)
library(MetBrewer)
library(RColorBrewer)
library(readxl)
```

```{r}
setwd("C:/Users/boloq/Box/Dissertation/FCAT/FCAT_data")
```


```{r}
ants_data <- read_excel("Ants_students_data_july2023.xlsx")
head(ants_data) #Primeiras linhas da tabela
summary(ants_data) #Sum?rio da tabela
str(ants_data) #Estrutura da tabela
names(ants_data) #Nomes das colunas da tabela
```
# Preparing the data frame 

Change the table form with species in columns
```{r}
#Just for transects 
ants_data_com <- dcast(ants_data, Site+Transect~Species, value.var = "Species")
View(ants_data_com)
```


We are going to  use the bait as a repetition so we going to creat a code of Transect+bait in a new colun
```{r}
ants_data$Transect_bait <- paste(ants_data$Transect,ants_data$Bait, ants_data$Method, sep="_")


#change the table form with species in columns
ants_data_com_Transect_bait <- dcast(ants_data, Site+Transect_bait~Species, value.var = "Count")
ants_data_com_Transect_bait <- ants_data_com_Transect_bait %>%
  replace(is.na(.),0) #Replace NAs with zeros.

```

# Creating a object only with the species
```{r}
com <- ants_data_com[,-c(1,2)] 
```

# Amostral coverge ====

Here we want to know how well we sample this comunity so we are more interested in the sites to do that  we creat a data.frame only with sites and with species in the columns.
```{r}
ants_data_com_site <-dcast(ants_data, Site~Species, value.var = "Species")
summary(ants_data_com_site)
```


#creating a object only with the species
```{r}
com.site <- ants_data_com_site[,-c(1)] 
com.site
```

Because ants are base on frequency we need to inclued the number of sample stations we had, inclucing a colomn with the number of samples
summary(ants_data_com_Transect_bait$Site)
```{r}
summary(ants_data_com_Transect_bait$Site)
com.site$n_samples<- c(15,17)

#give name to rows
row.names(com.site) <- ants_data_com_site[,1]

# Reorder to have n_samples in the first coloum
ncol(com.site)
com.site <- com.site[, c(19,1:18)]
com.site
```


# Species curve and rarefaction
We going to user the INEXT to do the sample curve and rarefaction
```{r}
out1 <- iNEXT(t(com.site), q=0, datatype="incidence_freq", endpoint = 20)
out1$DataInfo
out1$DataInfo$SC
#Number of species
(plot_sp <- ggiNEXT(out1, type=1, se=TRUE, grey=FALSE) 
  + theme_classic(base_size = 18) 
  + theme(legend.position="right")
  + labs(y="Number of species", x = "Number of Samples"))+
  scale_fill_discrete(labels = c("Reforestation", "Pasture"))+
  scale_shape_discrete(labels = c("Reforestation", "Pasture"))+
  scale_color_discrete(labels = c("Reforestation", "Pasture"))+ 
  scale_colour_manual(labels = c("Reforestation", "Pasture"), values= met.brewer("Lakota", n = , type = "discrete", override.order = TRUE)) +
  scale_fill_manual(labels = c("Reforestation", "Pasture"),values= met.brewer("Lakota", n = , type = "discrete", override.order = TRUE))

#ggsave(filename ="plot_sp.png" , plot = plot_sp, dpi=600, units=c("mm"), width=200, height=200)

#sample coverge 
(plot_C_amnostral <- ggiNEXT(out1, type=2, se=TRUE, grey=FALSE) 
  + theme_classic(base_size = 18) 
  + theme(legend.position="right")
  + labs(y="Sample coverage", x = "Number of Samples"))+
  scale_fill_discrete(labels = c("Reforestation", "Pasture"))+
  scale_shape_discrete(labels = c("Reforestation", "Pasture"))+
  scale_color_discrete(labels = c("Reforestation", "Pasture"))+ 
  scale_colour_manual(labels = c("Reforestation", "Pasture"), values= met.brewer("Lakota", n = , type = "discrete", override.order = TRUE)) +
  scale_fill_manual(labels = c("Reforestation", "Pasture"),values= met.brewer("Lakota", n = , type = "discrete", override.order = TRUE))


#ggsave(filename ="plot_cov.png" , plot = plot_C_amnostral, dpi=600, units=c("mm"), width=200, height=200)
```


# Diversity metrics
## Hill numbers

## Species richness
```{r}
(Richness <- sp_rich <- specnumber(com))
(tab_result <- data.frame(ants_data_com[,1:2], Richness))
aggregate(tab_result$Richness, list(tab_result$Site), FUN=mean)
#Shannon diversity
(tab_result$Shannon <- diversity(x = com, index = "shannon"))
aggregate(tab_result$Shannon, list(tab_result$Site), FUN=mean)
```

# Simpson diversity
```{r}
(tab_result$Simpson <- diversity(x = com, index = "simpson"))
aggregate(tab_result$Simpson, list(tab_result$Site), FUN=mean)
```

## Plots
```{r}
summary(tab_result)
tab_ggplot <- melt(tab_result, id= c("Site","Transect")) # crarte a table for the plot
tab_ggplot$Site<- as.factor(tab_ggplot$Site) #Sites as factors


cor_plot <- met.brewer("Lakota", n = , type = "discrete", override.order = TRUE) #choose colours

my_comparisons <- list("3", "Palacio")

diver<- ggplot(tab_ggplot, aes(x=Site, y=value, fill=Site)) +
  geom_boxplot()+
  theme_bw(base_size = 15)+
  scale_fill_manual(values = cor_plot) +
   scale_x_discrete(labels = c('Reforestation', 'Pasture')) +
  theme(legend.position =  "none")+ 
  labs(y="Diversity", x = "") +
  facet_wrap(~variable, scales = "free")+
  stat_compare_means(comparisons = NULL, method = "t.test", label = "p.format") 
diver

t.test(Richness ~ Site, data = tab_result)
t.test(Simpson ~ Site, data = tab_result)
t.test(Shannon ~ Site, data = tab_result)

#ggsave(filename ="diver.png" , plot = diver, dpi=600, units=c("mm"), width=200, height=200)
```

## Abundance
```{r}
abun <- ggplot(ants_data, aes(x=Site, y = Count, fill = Site)) +
  #geom_boxplot(aes(fill = Site, color = Site))
  geom_violin(aes(fill= Site, color = Site)) +
  theme_bw(base_size = 15) +
  scale_fill_manual(values = cor_plot) +
  scale_color_manual(values = cor_plot) +
  geom_point(color="black", alpha = 0.2,size =2, position = position_jitter(w=0.05))+
   scale_x_discrete(labels = c('Reforestation', 'Pasture')) +
  theme(legend.position =  "none")+ 
  labs(y="Abundance", x = "") +
  stat_compare_means(comparisons = NULL, method = "t.test", label = "p.format") 
abun

#ggsave(filename ="abundance.png" , plot = abun, dpi=600, units=c("mm"), width=200, height=200)

```


# Linear Models
```{r}
#Richness
mod1 <- glm(Richness~Site, poisson, data=tab_result)
(s <- summary(mod1))
anova(mod1, test="F")


#Shannon
mod2 <- lm(Shannon~Site, data=tab_result)
(s <- summary(mod2))
anova(mod2, test="F")


#Simpson
mod2 <- lm(Simpson~Site, data=tab_result)
(s <- summary(mod2))
anova(mod2, test="F")
```

## BETA DIVERSITY

# Turnover and Nestedness
```{r}
spp.pa <- decostand(com, method="pa")  # mudar para presencia e ausencia 
(jac <- beta.multi(spp.pa, index.family="jaccard")) # index.family="sorensen" - 
round(jac$beta.JTU / jac$beta.JAC, 3) # 89% of turnover/replacement of spp
round(jac$beta.JNE / jac$beta.JAC, 3) # 11% of nestedness of spp
1 - jac$beta.JAC # 18% similarity
```



# NMDS
# PERMANOVA



```{r}
ants_data_nmds <- dcast(ants_data, Site+Transect_bait~Species, value.var = "Count")
ants_data_nmds <- ants_data_nmds %>%
  replace(is.na(.),0) #Replace NAs with zeros.

summary(ants_data_nmds)

#creatging a object only with the species 
com.nmds <- ants_data_nmds[,-c(1,2)]

# Distance Matrix
ant.dist <- vegdist(t(com.nmds), method="bray", binary=FALSE)

x <- as.matrix(ant.dist)


# nmds <- metaMDS(ant.dist,  
#                 distance = "bray", 
#                 k = 2, 
#                 trymax = 999, 
#                 trace = F)
# 
# nmds
nmds <- metaMDS(x,
                distance = "bray",
                display = c("sites", "species"),
                noshare = TRUE,
                autotransform = FALSE,
                wascores = TRUE,
                tidy = TRUE,
                k = 2,trymax = 999)
stressplot(nmds)

summary(nmds)


par(mar=c(6,6,2,2))


sppscores(nmds) <- t(com.nmds) 

data.scores <- as_tibble(vegan::scores(nmds)$species)

NM.DS <- ggplot(data.scores, aes(NMDS1, NMDS2, color = ants_data_nmds$Site)) +
  geom_point(data = data.scores, aes(shape= ants_data_nmds$Site),size = 3, alpha = 0.5, stroke = 1) +
   scale_fill_discrete(labels = c("Reforestation", "Pasture"))+
  scale_shape_discrete(labels = c("Reforestation", "Pasture"))+
  scale_color_discrete(labels = c("Reforestation", "Pasture"))+ 
  scale_colour_manual(labels = c("Reforestation", "Pasture"), values= met.brewer("Lakota", n = , type = "discrete", override.order = TRUE)) +
  scale_fill_manual(labels = c("Reforestation", "Pasture"),values= met.brewer("Lakota", n = , type = "discrete", override.order = TRUE))+
  scale_fill_discrete(labels = c("Reforestation", "Pasture"))+
  geom_hline(yintercept = 0, colour = "grey50", size = 0.65) +
  geom_vline(xintercept = 0, colour = "grey50", size = 0.65) +
  theme_bw(base_size = 18) +
  theme(legend.position="right", 
       legend.title = element_text(face = "bold")) +
  guides(shape=guide_legend("Site"), color = guide_legend("Site"))+
  stat_ellipse(mapping = NULL, data = data.scores, geom = "path", size = 1,
position = "identity", type = "t", linetype = 1, level = 0.95, segments = 25,
na.rm = FALSE, show.legend = NA, inherit.aes = TRUE) 

NM.DS

#ggsave(filename ="nmds.png" , plot = NM.DS, dpi=600, units=c("mm"), width=200, height=200)

```

```{r}
#Pacote de paleta de cores 

#
cores <- brewer.pal(7, "Dark2")

plot(nmds, axes=F, cex.lab=3, cex.axis=3, xlab="", ylab="")
axis(side=1, lwd=4,cex.axis=1.3 )
axis(side=2, lwd=4, las=1.5, cex.axis=1.3)
box(lwd=3, bty="l")
mtext("NMDS1", side=1, line=3, cex=1.5)
mtext("NMDS2", side=2, line=3, cex=1.5)
mypch <- c(15,16,17,18,19,20,21)
mycol <- c(cores)
points(nmds, pch=mypch[as.numeric(ants_data_com$Site)], col=mycol[as.numeric(ants_data_com$Site)], cex=2)


ants_data_com 
#ggsave(filename ="Aim3_NMDS_06272022.jpeg" , plot = NM.DS, dpi=600, units=c("mm"), width=200, height=200)

#Plotando o elipse de tendencia 
summary(ants_data_com$Site)

treat=c(rep("Reforestation",15),rep("Pasture",17))
colors<- c(rep(cores[1],15),rep(cores[2],17))
#Plot convex hulls with colors baesd on treatment
for(i in unique(treat)) {
  ordiellipse(nmds$point[grep(i,treat),],draw="polygon",
              groups=treat[treat==i],col=colors[grep(i,treat)],label=F) }

###legand

legend(3, 2 , c("Reforestation", "Pasture"), pch=c(15,17), col=cores, cex=1.3, bty="n")
#text(4, 2.05,"Stress = 0.18")
```

# Functional Guild

```{r}
guild <- dcast(ants_data, Site+Transect~Guild, value.var = "Guild")

guild2 <- guild %>%
  pivot_longer(c(3:6), values_to = "Guild_count", names_to = "Guild") %>%
  mutate_at(vars(Site, Guild), as.factor) %>%
  dplyr::group_by(Site, Guild) %>%
  dplyr::summarise(guild_sum = sum(Guild_count))


# Plot
guiplot <- ggplot(guild2, aes(x= Site, y = guild_sum, fill = Guild)) +
  geom_col(aes(fill = Guild, color = Guild), position = "fill") +
  theme_bw(base_size = 15) +
  scale_fill_manual(values = cor_plot) +
  scale_color_manual(values = cor_plot) +
  scale_x_discrete(labels = c('Reforestation', 'Pasture')) + 
  theme(legend.position = "right")+ 
  labs(y="Relative Abundance (%)", x = "")
guiplot

#ggsave(filename ="guild.png" , plot = guiplot, dpi=600, units=c("mm"), width=200, height=200)

```

#ANOSIM - o anossim faz a mesma coisa mas calcuala a similaridade ao invez de analise de variancia
anosim(com.nmds, ants_data_nmds$Site, distance="bray",permutations = 1000 )
